### 2019-03-11 manuscript/introduction.tex
- cleanup of Intro, title alteration, couple of new refs

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility causes plants to reject their own pollen, and its loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}. % Gordon: Really? I find this hard to believe... %B:  sigh.
Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that unobserved states drive diversification patterns \citep{freyman_2018}.
Despite their apparent macroevolutionary role in diversification, breeding systems are clearly not the sole trait determining diversification rates.
Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
Moreover, polyploidy is not only a mere correlate of breeding system, but may directly cause the loss of self-incompatibility (SI) \citep{stout1942, lewis1947}, and facilitate the transition to selfing.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
This precludes the existence of SI polyploids in the most widespread system of SI, and creates a correlation between polyploidy and SC.
% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. 
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI. That's not a concern, I think, they are all "SI systems." And Dioecy etc. is not classically considered a "breeding system", but a "sexual system". That may be a part of the confusion. It's really not Gordon's problem, and we could do something to yet again re-define these terms? Emma?

Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
First, it may be of particular interest to establish their joint influence on lineage diversification.
It is possible that combinations of breeding system and ploidy do not result in predictable additive effects on speciation and extinction rates.
Second, the two traits may affect each other's state transitions, including their magnitude and ordering of transitions.
For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without in increase in ploidy, and then a subsequent diploidization. 
%Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems.

Two recent methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
%First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE). 
First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction. 
%Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}. 
Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with `hidden' or unspecified traits (HiSSE), which can affect the diversification process.
%The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen. 
Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables such variation to be explained by unobserved states, and reduces the elevated false-positive error rates \citep{rabosky_2015}.
%HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.  
%Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately. 
%Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
%However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
%Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates. 
A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates. 
%These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.

>>>>

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}. % Gordon: Really? I find this hard to believe... %B:  sigh.
Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2018}.
Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates.

Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
Polyploidy is not a mere correlate of transition in breeding system, but may directly cause the loss of SI \citep{stout1942, lewis1947}, as well as facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}.
Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
With exceedingly few exceptions \citep{hauck_2002,nunes_2006}, this generally precludes the existence of SI polyploids in the most widespread system of SI, which is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}.
This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date, and creates a strong correlation between polyploidy and SC \citep{robertson_2011}.
% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems. I don't think the de Vos paper is dealing with SI systems. If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying. 
% R- Boris what do you think is the best way to proceed here? Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI. That's not a concern, I think, they are all "SI systems." And Dioecy etc. is not classically considered a "breeding system", but a "sexual system". That may be a part of the confusion. It's really not Gordon's problem, and we could do something to yet again re-define these terms? Emma?

Given that changes in ploidy and breeding systems are causally related, and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
First, it may be of interest to establish their joint influence on lineage diversification.
Combinations of breeding system and ploidy may not have individually predicted, additive effects on speciation and extinction rates.
Second, the two traits may affect each other's state transitions, including both magnitudes and ordering.
For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without in increase in ploidy, and then a subsequent diploidization. 
%Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems.

Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
%First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE). 
First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction. 
%Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}. 
Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
%The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen. 
Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and reduces the elevated false-positive error rates \citep{rabosky_2015}.
%B: This may be a good portion of Discussion--faults and qualities of methods?
%HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.  
%Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately. 
%Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood. 
%However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
%Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates. 
A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates. 
%These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.

### 2019-03-11 manuscript/main.tex
First part of the discussion completed, missing some of it
- outline in discussion, removed errant mark in main
fixing a plot
- cleanup of Intro, title alteration, couple of new refs

% Accumulate ideas for titles here...
%
% When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae % E: Fun, but I think we need something more formal
%
%% When the favorite trait is not enough: the link of polyploidy and breeding system in diversification %B: <- eh?
%
% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
%
% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
%
%B: associations isn't exactly right; I think it somewhat undermines the causality from DP to IC; anyway, it's not particularly bad.
%
%% Interaction of ploidy and breeding system in lineage diversification %B: <- how 'bout that?
%
\begin{center}
    \textsc{Interacting associations between ploidy, breeding system, and lineage diversification}

@@

\noindent 
Rosana Zenil-Ferguson$^{1,\dagger}$,
%
J.~Gordon Burleigh$^{2}$,
%
William A.~Freyman$^{3}$,
%
Boris Igi\'c$^{4}$,
%
Itay Mayrose$^{5}$,
%
and Emma E.~Goldberg$^{3}$
%
\end{center}
%
\noindent$^{2}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
%
\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
%
\noindent$^{3}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
%
\noindent$^{4}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
%
\noindent$^{5}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
%
\noindent$^\dagger$\textit{Author for correspondence:}.\\

>>>>

% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
%
% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
%
% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
%
\begin{center}
    \textsc{Interaction between ploidy, breeding system, and lineage diversification}

@@

\noindent 
Rosana Zenil-Ferguson$^{1,\dagger}$,
%
J.~Gordon Burleigh$^{2}$,
%
William A.~Freyman$^{3}$,
%
Boris Igi\'c$^{4}$,
%
Itay Mayrose$^{5}$,
%
and Emma E.~Goldberg$^{3}$
%
\end{center}
%
\noindent$^{1}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
%
\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
%
\noindent$^{3}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
%
\noindent$^{4}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
%
\noindent$^{5}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
%
\noindent$^\dagger$\textit{Author for correspondence}.\\
